---
title: "Wrangling Data"
author: "Andy"
date: "16 May 2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data Wrangling Webinar

This is a step-through of the work done in the Data Wrangling Webinar from RStudio.

Firstly, load essential packages:

```{r}
library("tidyr")
library("dplyr")
```


This is an example of tidy data. Each column is a variable to be called upon.

```{r}
load("storms.rdata")
storms
```

Whilst this is an examploe of data that is not yet "tidy"

```{r}
load("cases.rdata")
cases
```

Converting it to a "tidy" data frame:

```{r}
gather(cases, "year","n",2:4)

```

Within the gather function:  
 --- **cases** is the data frame to be transformed  
 --- **year** is the name of the new column variable, representing the column names of the original data frame  
 --- **n** is the name of the new column variable that includes all the values of the data frame  
 --- **2:4** is used to show that only columns 2 to 4 of the original data frame need to be transformed into our new data frame.