---
title: "Wrangling Data"
author: "Andy"
date: "16 May 2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data Wrangling Webinar

This is a step-through of the work done in the Data Wrangling Webinar from RStudio.

Webinar is here: <https://rstudio.com/resources/webinars/data-wrangling-with-r-and-rstudio/>

Firstly, load essential packages:

```{r}
library("tidyr")
library("dplyr")
```


This is an example of tidy data. Each column is a variable to be called upon.

```{r}
load("storms.rdata")
storms
```

Whilst this is an examploe of data that is not yet "tidy"

```{r}
load("cases.rdata")
cases
```

The issue is that the same variable with multiple years is split across the data frame. Really we want a 3 column dataframe - Country / Year / "n"

Converting it to a "tidy" data frame we use the *gather* function:

```{r}
gather(cases, "year","n",2:4)

```

Within the gather function:  
 --- **cases** is the data frame to be transformed  
 --- **year** is the name of the new column variable, representing the column names of the original data frame  
 --- **n** is the name of the new column variable that includes all the values of the data frame  
 --- **2:4** is used to show that only columns 2 to 4 of the original data frame need to be transformed into our new data frame.
 
 Next, to move onto different issues with another dataset.
 
```{r}
load("pollution.rdata")
pollution
```
Here, the issue is that there is a duplication of variables in the "city" column which are matched to the size. We want to separate this out into 3 columns of city / large / small. In this way the data can be called easily with reference to a variable in the data frame.

Here we use the *spread* function to make it "tidy"

```{r}
spread(pollution, size, amount)
```

The spread function inputs are:  
---**pollution** as the data frame to transform  
---**size** as the key column to use with each factor used as a new column  
---**amount** as the associated values to be used in the dataframe  

Note that *spread* and *gather* are opposites. Applying both should return the original data frame. 

Another useful function is *separate* which creates additional columns from one original column:

```{r}
storms2<-separate(storms, date, c("year","month", "day"), sep = "-")
storms2
```

The individual parts of separate are fairly 

*unite* is the opposite of *separate*

```{r}
storms3<-unite(storms2, "date", year, month, day, sep = "-")
storms3
```

